The Ranked Nearest Neighbour Interchange space

### 2019-05-03 paper.tex
how to use Bubble Sort

%Do we restrict everything to RNNI? If not, make sure that we correctly distinguish between RNNI and RNNIu (for instance in the algorithm section)

@@

\section{Approximating Distances}

>>>>

%Do we restrict everything to RNNI? Make sure that we correctly distinguish between RNNI and RNNIu (for instance in the algorithm section)

@@

\section{Algorithms}

We are now going to introduce different algorithms that computes paths in $\rnni$.
\cite{Gavryushkin2017} gave an algorithm for computing the whole $\rnni$ space.
We implemented this algorithm and are able to use it on trees with up to $8$ taxa.
Therefore, we can compare the paths given by the algorithms presented in this chapter with the exact shortest path for trees with up to $8$ taxa.

% Shall we mention the data structure we use (list of sets) - this might be useful to see the connection to the partition lattice.

A modified version of the classic Bubble Sort \todo{cite} algorithm returns a caterpillar path that is shortest among all caterpillar paths between given caterpillar trees $T_1, T_2$.
Let $T_1 = (\ldots((1,2)3) \ldots )n), T_2 = (\ldots((\sigma_1,\sigma_2)\sigma_3) \ldots )\sigma_n)$ be two caterpillar trees.
They can be seen as sequences $(1,2,\ldots,n)$ and $(\sigma_1, \sigma_2, \ldots, \sigma_n)$ that need to be sorted.
$\nni$ moves between caterpillar trees correspond to exchanging two neighboured elements in a sequence.
The only exception of this are $\nni$ moves that change the cherry of a tree.
The corresponding moves on sequences are exchanged of either the second and third element or the first and third element.
On the other hand, there is no difference between two the trees corresponding to two sequences where only the first two taxa are exchanged.
Since all pairs of taxa whose order in $T_1$ is different from the one in $T_2$ need to be exchanged by $\nni$ moves on a caterpillar path, Bubble Sort will give a shortest path among all caterpillar paths.
Just the moves for the first three positions of the sequence must be adapted.
Let us consider the first three elements $(a,b,c,\ldots)$ of a sequence representing a caterpillar tree on a caterpillar path between $T_1$ and $T_2$
If the order of these elements is the same in $T_2$, there are no swaps needed for these taxa.
If just one of the two taxa $a$ and $b$ is on the other side of $c$ in $T_2$, this one will change wih $c$.
But if both $a$ and $b$ are on the other side of $c$ in $T_2$, the one which is the righmost of all three taxa in $T_2$ swaps with $c$ at first.
This makes sure that there is no additional swap of $a$ and $b$ needed on the path to $T_2$.
\todo{This paragraph is just a draft. Is it needed at all?}
